---
title: "Myungje_Kim_Final_Project"
author: "Myungje Kim"
format: 
  html:
    embed-resources: true
editor: visual
execute:
  echo: true
  message: false
  warning: false
---

```{r}
library(tidyverse)
library(openintro)
library(gridExtra)
```

```{r}
data("birds")
```

This data has phase of flight and time of day as factors. birds adj 2 and 3 is just to show that 16671 has the same observations when I filter it with NAs one by one.

```{r}
birds_all_states <- birds |>
  select(phase_of_flt, time_of_day, state) |>
  na.omit()

birds_CA <- birds |>
  select(phase_of_flt, time_of_day, state) |>
  na.omit() |>
  filter(state == "CA") |>
  mutate(
    phase_of_flt = as.factor(phase_of_flt),
    time_of_day = as.factor(time_of_day)
  )

birds_CA

birds_Otherst <- birds |>
  select(phase_of_flt, time_of_day, state) |>
  na.omit() |>
  filter(state != "CA") |>
  mutate(
    phase_of_flt = as.factor(phase_of_flt),
    time_of_day = as.factor(time_of_day)
  )

nrow(birds_CA) + nrow(birds_Otherst) == nrow(birds_all_states)

#birds_adj2 <- birds |>
#  filter(!is.na(phase_of_flt)) |>
#  mutate(
#    phase_of_flt = as.factor(phase_of_flt)
#  )
         

#birds_adj3 <- birds_adj2 |>
#  filter(!is.na(time_of_day)) |>
#  mutate(
#    time_of_day = as.factor(time_of_day)
#  )


#birds_adj2
#unique(birds_adj2$phase_of_flt)

#birds_adj3
#unique(birds_adj2$time_of_day)
```

Checking if there is NA values in each dataset.

```{r}
#birds |>
#  select(state) |>
#  count(state == "NA")

#birds_CA |>
#  select(state) |>
#  count(state == "NA")

#birds_Otherst |>
#  select(state) |>
#  count(state == "NA")

```

# Data filtering

```{r}
plot1_data <- birds |>
  select(phase_of_flt, state) |>
  filter(!is.na(phase_of_flt), !is.na(state)) |>
           mutate(state1 = ifelse(state == "CA", "CA", "Other States"))

plot2_data <- birds |>
  select(phase_of_flt, time_of_day, state) |>
  filter(!is.na(phase_of_flt), !is.na(state), !is.na(time_of_day)) |>
  mutate(state2 = if_else(state == "CA", "CA", "Other States"))
```

# Plot 1

Version 1

```{r}
ggplot(plot1_data,
       aes(x = fct_rev(fct_infreq(phase_of_flt)),
           fill = state1)) +
  geom_bar() +
  coord_flip() +
  labs(title = "Collision Frequency vs Phase Of Flight ",
       x = "Frequency(Collision)",
       y= "Phase of Flight",
       fill = NULL)

```

Version 2

```{r}
p1 <- ggplot(birds_CA ,aes(x = fct_rev(fct_infreq(phase_of_flt)), fill = phase_of_flt)) +
  geom_bar(show.legend = FALSE) +
  coord_flip() +
  labs(title = " Frequency by Phase: CA",
       y = "Collision frequency",
       x = "Phase of Flight")

p2 <- ggplot(birds_Otherst ,aes(x = fct_rev(fct_infreq(phase_of_flt)), fill = phase_of_flt)) +
  geom_bar(show.legend = FALSE) +
  coord_flip() +
  labs(title = "Frequency by Phase: Other States",
       y = "Collision frequency",
       x = "Phase of Flight")

grid.arrange(p1, p2, nrow = 1)
```

This plot compares California and other states in terms of which **flight phase** had the most bird collisions.\
In both groups, most collisions happen during the **approach** and **landing** phases. This is expected because planes are flying lower and slower then, which increases the chance to hit birds.

In **California**, the collisions are more evenly spread across all phases, while in **other states**, they are mainly in one or two phases, especially during **approach**.\
Even if the scales are different, the trend is the same: the **approach phase** has the most collisions in both cases.

# Plot2

```{r}
p3 <- plot2_data |>
  filter(state2 == "CA") |>
  ggplot(aes(x = fct_infreq(phase_of_flt), fill = time_of_day)) +
  geom_bar(position = position_dodge(width = 0.8), width = 0.7) +
  facet_wrap(~state2, ncol = 2) +
  labs(title = "Phase of Flight vs Time of Day (CA)",
       x = "Phase Of Flight",
       y = "Time of Day")
```

```{r}
p4 <- plot2_data |>
  filter(state2 != "CA") |>
  ggplot(aes(x = fct_infreq(phase_of_flt), fill = time_of_day)) +
  geom_bar(position = position_dodge(width = 0.8), width = 0.7) +
  facet_wrap(~state2, ncol = 2) + 
  labs(title = "Phase of Flight vs Time of Day (Other States)",
       x = "Phase Of Flight",
       y = "Time of Day")

```

This plot shows how **time of day** affects collisions for California and other states.\
In both places, most collisions happen during the **daytime**, and there are fewer at **night** or **dawn/dusk**.

In **California**, the collisions during the day are more spread across different flight phases. But in **other states**, they are mostly concentrated in the **approach phase** during the day.\
So even though the number of collisions is not on the same scale, the pattern is clear: **daytime and approach phase** are the most common times for bird collisions in both California and other states.

```{r}
grid.arrange(p3, p4, nrow = 1)
```
